Adding Sensitivity to 21cm Inteferometric Probes of Reionization by Optimizing Choice of Baselines

### 2017-03-17 Rotation paper/draft_working.tex
save

\begin{abstract}
We introduce an accurate method to extract additional sensitivity from data of redundant radio arrays. Our method efficiently finds optimal baselines to cross correlate, including baselines that are slightly different in length and orientation. 
Using the latest and final 128-element version of Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as the planned 37, 128, 240 and 350 element versions of the Hydrogen Epoch of Reionization Array (HERA) now under development, we illustrate how our method applies to different array configurations and predict the sensitivity improvements of including each baseline pairs. The extremely natural and simple implementation of this method
would increase the sensitivity of PAPER-128 by up to [~50] percent, and increase the out-rigger imaging sensitivity of HERA-350 by a factor of 2. 
\end{abstract}

@@

The epoch of reionization, or cosmic dawn represents the last key
stage of the universe's early evolution. Study of this event stands at
the intersection of cosmology and astrophysics. Understanding this
event is important not only because it is a scientific goal
of its own, but also because it can provide crucial information
to fundamental nature of inflation, neutrino mass and phenomenology
of the first stars and galaxies (e.g. \cite{LiuOpticalDepth}).

Arguably the most promising observational probe of the epoch of reionization
comes from measurement of the ``spin-flip'' transition of neutral
hydrogen of characteristic wavelength 21cm \cite{Furlanetto2006181,PritchardLoeb}.
While probes such as quasar spectra rely on emission and scattering
of free electrons, and are thus limited to the lower redshifts after
the completion of reionization, the 21cm line directly probes the
abundance and distribution of neutral hydrogen, and thus would potentially
shed light on all stages of reionization. Radio interferometry efforts
to measure the 21cm power spectrum has been a top priority in recent years of astronomy.
Current generation instruments include the Precision Array for Probing
the Epoch of Reionization (PAPER) \cite{Ali2015,paper32}, Murchison
Widefield Array (MWA) \cite{MWA}, Low Frequency Array (LOFAR) \cite{LOFAR},
and next generation instruments such as the Hydrogen Epoch of Reionization
Array (HERA) (e.g. \cite{HERA,HERAconfiguration,HERABEAM1,HERADISH2})
and the Square Kilometer Array Low (SKA low) (e.g. \cite{SKA1}) are
under construction.

One of the main challenges of interferometric observations of cosmic
21cm line is the foreground contamination. Both sources within our
galaxy, and to a lesser distant sources outside of our galaxy emit
radio contaminations (via for example synchrotron processes) up to five orders of magnitude stronger than the
reionization signal. There are two common methods to deal with the
foreground contamination. In the first, sometimes called the foreground
removal technique, individual sources are identified and removed in the image domain.
The other technique, commonly referred to as 'foreground avoidance',
makes use of the fact that most common foreground contaminations have a smooth
spectrum, and thus is constrained to a Fourier domain ``foreground
wedge'' \cite{wedge1,wedge2}. Thus contamination can be avoided
by restricting our observation to ``outside the wedge'', where the data is currently noise dominated. The main
challenge of using the avoidance technique is thus the sensitivity, or
the signal to noise ratio. This is the motivation for the design of
the maximum redundancy arrays such as PAPER and HERA. While traditional
imaging arrays focus on Fourier mode uv coverage, the redundant arrays
focus on redundant measurements of the same Fourier mode
with many tightly packed antennas at equal spacing. Since baselines
of the same length and orientation measure the same Fourier modes
on the sky, a maximum redundancy array is able to increase the
signal to noise ratio by averaging over measurements of the same baseline.
\cite{Ali2015} provides the newest upper limit to the power spectrum
measurements with the PAPER-64 data. From some models we roughly expect
that the sensitivity required for detection is a factor of 10 away.
This paper explores a new technique to add sensitivity to the analysis
of redundant arrays, by making use of the earth's rotation. This technique
can be applied to redundant arrays such as PAPER and HERA to increase
the sensitivity by up to a factor of [1.5].

The earth's rotation causes the baselines to pick up different modes of the sky with time. 
This effect is used extensively to improve uv coverage in imaging with minimum redundancy arrays. In
a maximally redundant array, baselines that are slightly different
rotate into each other at a time delay. Here we present a method to
improve the sensitivity of the maximum redundancy arrays by including
the uv redundancy of different baselines at a time lag. Recently the authors of \cite{wterm} proposed an algorithmic search approach to extract information from all near-equivalent baselines. We here propose a simple and elegant theoretical formalism that in addition allows us to identify good baselines to cross-correlate and predict the weight of sensitivity with speed and accuracy.

Baselines of the same length and orientation are traditionally called
``redundant baselines'', because they measure the same Fourier mode
in the sky. In order to eliminate confusion and ambiguity, we shall
introduce slightly different terminology. We'll call baselines that
are the same length and orientation ``equivalent baselines'', inspired
by the mathematical notion of equivalency classes. Two equivalent
baselines will be redundant with each other simultaneously at all
times. Non-equivalent baselines can also be partially redundant if their
respected time series is shifted with respect to one another by a
certain delay. In other words, one baseline can be ``rotated into''
another. We call these ``near-equivalent baselines''. It is our
goal to a) identify the near-equivalent baselines that give good
redundancy, b) to find the optimal time offset for a given pair of baselines, and
c) to quantify the sensitivity improvement associated with cross multiplying
such a pair of near-equivalent baselines. In section
2 we explain the theoretical basis for this cross multiplication.
In section 3 we present details of the numerical implementation of
this technique. In section 4 we discuss the expected sensitivity improvement
with this method for HERA and PAPER-128 and with section 5 we conclude.

@@

We shall use the 128-element PAPER array to demonstrate. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels are show
in Fig. \ref{fig:AntPos}. We see the antenna spacing in North-South
directions are comparatively close (4m), so that baselines such as
0\_44 and 0\_7 are very close to equivalent. In the bottom panel,
the two baselines labeled with the red arrow are equivalent, with
separation denoted by sep1,0, for the antennas are separated by 1
unit east and 0 unit north. Similarly, the baselines labeled with
the yellow and green arrows are examples of sep1,1 and sep-1,1, respectively.
Note sep1,0 and sep-1,0 for example are the same baselines and should
not be counted twice. Antennas in purely north-south baselines
are close enough to induce cross talks, and hence are not suitable
for use. The original PAPER-64 analysis used three classes of baselines, here
equivalent to 
sep2,0, sep2,1 and sep-2,1 \cite{Ali2015} of PAPER-128. There each of these classes
of baselines are cross multiplied to itself. We shall see that in addition these
baseline classes can be cross multiplied with a time offset.

Given a point source on the sky, each baseline maps the
source to a point in uv plane. As the earth rotates with respect to
the source, the point traces out tracks in the uv-plane. 
The uv tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith, are shown in Fig. \ref{fig:Tracks}.

@@

Roughly speaking, we can identify
redundancy of near-equivalent baselines as crossings
of the uv tracks. Although this is a valid quick method to identify some baselines to cross-multiply, we find
that it is not accurate enough for time offset determination. The reason is, for wide beam antennas such as PAPER, the sources are not point-like. 
In other words, when tracks of
one point of the sky are crossing in uv-plane, those corresponding to another is not. To accurately
constrain redundancy we must take into account of finite beam width. A more rigorous treatment is thus required. 
We show sample beams of HERA and PAPER antennas in Fig.\ref{fig:Beam} for reference.

@@

We take the visibility as commonly defined in the literature (e.g.
\cite{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(b) & =\int d\Omega A_{\nu}(\hat{s})\phi(\nu)I_{\nu}(\hat{s})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{s}\right],\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{s})\phi(\nu)T(\hat{s})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{s}\right],
\end{aligned}
\label{eq:Vis1}
\end{equation}
Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. $A_{\nu}$ is the (frequency
dependent) beam, and $I$ is the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. Note that $A$ here is the baseline beam, i.e. product of two antenna beams, examples of which are shown in Fig.\ref{fig:Beam}. $\phi(\nu)$ is the frequency band profile. For PAPER, for example,
the bands studied are roughly a top-hat from 144MHz to 154 MHz.

We define the delay transformed visibility \cite{delay-transform}:
\small
\begin{equation}
\begin{aligned}V(b,\tau) & =\int d\nu V_{\nu}(b)\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{s},\nu)\phi(\nu)T(\hat{s},\nu)\exp\left[-2\pi i\nu\left(\frac{b\cdot\hat{s}}{c}+\tau\right)\right]
\end{aligned}
.\label{eq:Vb1}
\end{equation}
\normalsize

Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observation coordinates $\hat{s}$ and $\nu$. Ultimately,
we would like to relate the data, collected with coordinates $\hat{s}$
and $\nu$, to the power spectrum, written with cosmological coordinates
$\boldsymbol{r}$ and $\boldsymbol{k}$. We start by noticing that
{[}Gerry: change all r and k into bold{]} 
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
 & \equiv D_{c}-Y\Delta\nu,
\end{aligned}
\]
where $\nu_{21}=1420MHz$ is the 21cm transition rest frequency, $\nu_{0}$
a reference central frequency with corresponding redshift $z_{0}$,
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{D_{c}-r}{Y}+\nu_{21}.\label{eq:nur}
\end{equation}
Thus $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$.

We can thus rewrite the delayed transformed visibility as 
\small
\[
\begin{aligned}V(b,\tau) & =\frac{1}{X^{2}Y}\int_{H}d^{3}rA(\vec{r})\phi(r)I(\vec{r})\exp\left[-2\pi i\left(\frac{b}{c}\cdot\hat{r}+\tau\right)\nu_{r}\right]\end{aligned}.
\]
\normalsize
Here $(r_{x},r_{y},r_{z})=(X\hat{s}_{x},X\hat{s}_{y},Y\nu)$, and
$(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\eta)$ relate
the cosmological coordinates $r$ and $k$ to the measured coordinates.
We have written $\nu_{r}$ to remind us that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}. The beam reception pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines.

With a time offset, the beam pattern has moved relative to the sky.
Here we choose to fix the sky, and denote the rotated coordinates
of the beam pattern with the operator $\Gamma$. With implicit bounds
of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} & \langle V^{*}(b,\tau)V(b',\tau')\rangle\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(r)T(r')\rangle\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\kappa)T(\kappa')\rangle e^{-i(\kappa r-\kappa'r')}\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\kappa(r-r')}\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\xi(r-r')A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
 & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(r-r')A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(r)A(\Gamma r)||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{r}\cdot\frac{b}{c}-\hat{\Gamma r}\cdot\frac{b'}{c}\right)\right],\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{s},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{s}\cdot\frac{b}{c}-\hat{\Gamma s}\cdot\frac{b'}{c}\right)\right],
\end{aligned}
\label{eq:main}
\end{equation}

where in transition from cosmological coordinates back to observing coordinates we have written $\hat{r}\equiv\hat{s}$, and 
\begin{equation}
\Phi_{b,\tau}(r,\Gamma r')=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(b\cdot\nu_{r}\hat{r}-b'\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
\end{equation}
\end{widetext}
The second to third line of Eq.(\ref{eq:main}) follows from assumption of Gaussian random
sky, and the third to fourth line follows from the assumption that
the 3D power spectrum varies negligibly over the region of interest
so that $\hat{P_{21}}(k+k_{2})\approx\hat{P_{21}}(k)$. Since $\Gamma$
is a sky rotation, it doesn't affect $\nu$, hence we have taken $\nu_{r}$
outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end. This is significant because this means that correlation is peaked at the same time-offset for all delay channels.

Finally, since the beam pattern and bandwidth are given in $\hat{s}$
and $\nu$, we convert the integral back to these coordinates to get
the general relation between the delay-transformed visibilities and
the power spectrum:

\begin{equation}
\begin{aligned} & \langle V^{*}(b,\tau)V(b',\tau')\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{s},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\\
 & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{s}\cdot\frac{b}{c}-\hat{\Gamma s}\cdot\frac{b'}{c}\right)\right].\end{aligned}
\label{eq:final}
\end{equation}
So roughly speaking the cross multiplications of visibilities at a time delay
in uv-space is proportional to the power spectrum times the Fourier
transform of the cross multiplied beam pattern. We can then in principle
combine information from different baseline pairs if we correct for
the phase and normalization. As a check, when applied to equivalent baselines,
$b=b'$, $\hat{s}=\hat{\Gamma s}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}.

If this result is robust, then we can achieve all our goals stated in the introduction, i.e. to identify 
candidate baseline pairs, to find the optimal time offset, and to quantify the sensibility, by computing the integral
in Eq.(\ref{eq:final}) for various time offsets. Thus we shall first test its robustness numerically.

\subsection{Rephasing \label{sec:rephs}}
Before we discuss tests, there is one subtlety when applying Eq. \ref{eq:final}. The optimal time offset is given by the $\Gamma$ that maximizes the integral. However, first integrating over the spatial dimension $\Omega$, we see the phase term at peak time-offset is inevitably frequency dependent. This frequency dependence of the phase would lead to destructive interference when we integrate over frequency.

>>>>

\begin{abstract}
We introduce an accurate method to extract additional sensitivity from data of redundant radio arrays. Our method efficiently finds optimal baselines to cross correlate, including baselines that are slightly different in length and orientation. 
Using the latest and final 128-element version of Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as the planned 37, 128, 240 and 350 element versions of the Hydrogen Epoch of Reionization Array (HERA) now under development, we illustrate how our method applies to different array configurations and predict the sensitivity improvements of including each baseline pairs. The extremely natural and simple implementation of this method
would increase the sensitivity of PAPER-128 by about to $10\%$, and increase the out-rigger imaging sensitivity of HERA-350 by a factor of 2. 
\end{abstract}

@@

The epoch of reionization, or cosmic dawn represents the last key
stage of the universe's early evolution. Study of this event stands at
the intersection of cosmology and astrophysics. Understanding this
event is important not only because it is a scientific goal
of its own, but also because it can provide crucial information
to fundamental physics of inflation, neutrino mass and phenomenology
of the first stars and galaxies (e.g. \cite{LiuOpticalDepth}).

Arguably the most promising observational probe of the epoch of reionization
comes from measurement of the ``spin-flip'' transition of neutral
hydrogen of characteristic wavelength 21cm \cite{Furlanetto2006181,PritchardLoeb}.
While probes such as quasar spectra rely on emission and scattering
of free electrons, and are thus limited to the lower redshifts towards the end of reionization, the 21cm line directly probes the
abundance and distribution of neutral hydrogen, and thus would potentially
shed light on all stages of reionization. Radio interferometry efforts
to measure the 21cm power spectrum has been a top priority in recent years of astronomy.
Current generation instruments include the Precision Array for Probing
the Epoch of Reionization (PAPER) \cite{Ali2015,paper32}, Murchison
Widefield Array (MWA) \cite{MWA}, Low Frequency Array (LOFAR) \cite{LOFAR}. Next generation instruments such as the Hydrogen Epoch of Reionization
Array (HERA) (e.g. \cite{HERA,HERAconfiguration,HERABEAM1,HERADISH2})
and the Square Kilometer Array Low (SKA low) (e.g. \cite{SKA1}) are
in planning and construction.

One of the main challenges of interferometric observations of cosmic
21cm line is foreground contamination. Both sources within our
galaxy, and to a lesser distant sources outside of our galaxy emit
radio contaminations (via for example synchrotron processes) up to five orders of magnitude stronger than the
reionization signal. There are two common methods to deal with the
foreground contamination. In the first, sometimes called the foreground
removal technique, individual sources are identified and removed in the image domain.
The other technique, commonly referred to as 'foreground avoidance',
makes use of the fact that most common foreground contaminations have a smooth
spectrum, and thus is constrained to a Fourier domain ``foreground
wedge'' \cite{wedge1,wedge2}. Thus contamination can be avoided
by restricting our observation to ``outside the wedge'', where the data is currently noise dominated. The main
challenge of using the avoidance technique is thus the sensitivity, or
the signal to noise ratio. This is the motivation for the design of
the maximum redundancy arrays such as PAPER and HERA. Traditional
imaging arrays are designed to image localized bright sources and thus favor non-redundant patterns for uv coverage. The redundant arrays on the other hand
focus on high-sensitivity measurements of the same Fourier mode
with many tightly packed antennas at equal spacing. Since baselines
of the same length and orientation measure the same Fourier modes
on the sky, a maximum redundancy array is able to increase the
signal to noise ratio by averaging over measurements of the same baseline.
\cite{Ali2015} provides the newest upper limit to the power spectrum
measurements with the 64-element version of PAPER(henceforth referred to as PAPER-64). From some models we roughly expect
that the sensitivity required for detection is a factor of 10 away.
This paper explores a new technique to add sensitivity to the analysis
of redundant arrays, by making use of the earth's rotation. This technique
can be applied to redundant arrays such as PAPER and HERA to increase
the sensitivity by up to a factor of [1.5].

The earth's rotation causes the baselines to pick up different modes of the sky with time. 
This effect is used extensively to improve uv coverage in imaging with minimum redundancy arrays. In
a maximally redundant array, baselines that are slightly different
rotate into each other at a time delay. Here we present a method to
improve the sensitivity of the maximum redundancy arrays by including
the uv redundancy of different baselines at a time lag. Recently the authors of \cite{wterm} proposed an algorithmic search approach to extract information from all near-equivalent baselines. We here present a simple and elegant theoretical formalism that in addition allows us to identify good baselines to cross-correlate and predict the weight of sensitivity with speed and accuracy.

Baselines of the same length and orientation are traditionally called
``redundant baselines'', because they measure the same Fourier mode
in the sky. In order to eliminate confusion and ambiguity, we shall
introduce slightly different terminology. We'll call baselines that
are the same length and orientation ``equivalent baselines'', inspired
by the mathematical notion of equivalency classes. Two equivalent
baselines will be redundant with each other simultaneously at all
times. Non-equivalent baselines can also be partially redundant if their
respected time series is shifted with respect to one another by a
certain delay. In other words, one baseline can be ``rotated into''
another. We call these ``near-equivalent baselines''. It is our
goal to a) identify the near-equivalent baselines that give good
redundancy, b) to find the optimal time offset for a given pair of baselines, and
c) to quantify the sensitivity improvement associated with cross multiplying
such a pair of near-equivalent baselines. In section
2 we explain the theoretical basis for this cross multiplication.
In section 3 we present numerical tests of
this technique as well as the expected sensitivity improvement
with this method for HERA and PAPER-128 and with section 4 we conclude.

@@

We shall use the 128-element PAPER array to demonstrate. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels are show
in Fig. \ref{fig:AntPos}. We see the antenna spacing in North-South
directions are comparatively close (4m), so that baselines such as
0\_44 and 0\_7 are very close to equivalent. In the bottom panel,
the two baselines labeled with the red arrow are equivalent, with
separation denoted by sep1,0, for the antennas are separated by 1
unit east and 0 unit north. Similarly, the baselines labeled with
the yellow and green arrows are examples of sep1,1 and sep-1,1, respectively.
Note sep1,0 and sep-1,0 for example are the same baselines and should
not be counted twice. Antennas in purely north-south baselines
are close enough to induce cross talks, and hence are not suitable
for use. The original PAPER-64 analysis (\cite{Ali2015}) used three classes of baselines, here
equivalent to 
sep2,0, sep2,1 and sep-2,1 \cite{Ali2015} of PAPER-128. There each of these classes
of baselines are cross multiplied to itself. We shall see that in addition these
baseline classes can be cross multiplied with a time offset.

Given a point source on the sky, each baseline maps the
source to a point in uv plane. As the earth rotates with respect to
the source, the point traces out tracks in the uv-plane. 
We show in Fig. \ref{fig:Tracks} uv tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith.

@@

Roughly speaking, we can identify
redundancy of near-equivalent baselines as crossings
of the uv tracks. Although this is a valid quick method to identify some baselines to cross-multiply, we find
that it is not accurate enough for time offset determination. The reason is that UV tracks are traces of a point source, which though adequate for traditional narrow beam imaging, are not applicable for wide-beam receivers like that of PAPER.
Here the uv-imprint of the sky for a given baseline and frequency is really a cloud that's the track convolved with the appropriate point-spread function. To accurately
constrain redundancy, a more rigorous treatment is thus required. 
We show sample beams of HERA and PAPER antennas in Fig.\ref{fig:Beam} for reference.

@@

We take the visibility as commonly defined in the literature (e.g.
\cite{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],
\end{aligned}
\label{eq:Vis1}
\end{equation}
Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. $A_{\nu}$ is the (frequency
dependent) beam, and $I$ is the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. Note that $A$ here is the baseline beam, i.e. product of two antenna beams, examples of which are shown in Fig.\ref{fig:Beam}. $\phi(\nu)$ is the frequency bandpass profile. In practice,
power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).

We define the delay transformed visibility \cite{delay-transform}:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]
\end{aligned}
.\label{eq:Vb1}
\end{equation}
\normalsize

Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observation coordinates $\hat{\boldsymbol{s}}$ and $\nu$. Ultimately,
we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
and $\nu$, to the power spectrum, written with cosmological coordinates
$\boldsymbol{r}$ and $\boldsymbol{k}$. We start by noticing that
{[}Gerry: change all r and k into bold{]} 
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
 & \equiv D_{c}-Y\Delta\nu,
\end{aligned}
\]
where $\nu_{21}=1420MHz$ is the 21cm transition rest frequency, $\nu_{0}$
a reference central frequency with corresponding redshift $z_{0}$,
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{D_{c}-r}{Y}+\nu_{21}.\label{eq:nur}
\end{equation}
Thus $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$.

We can thus rewrite the delayed transformed visibility as 
\small
\[
\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{1}{X^{2}Y}\int_{H}d^{3}rA(\boldsymbol{r})\phi(r)I(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned}.
\]
\normalsize
Here $(r_{x},r_{y},r_{z})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y},Y\nu)$, and
$(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relate
the cosmological coordinates $r$ and $k$ to the measured coordinates.
We have written $\nu_{r}$ to remind us that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}. The beam reception pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines.

With a time offset, the beam pattern has moved relative to the sky.
Here we choose to fix the sky, and denote the rotated coordinates
of the beam pattern with the operator $\Gamma$. With implicit bounds
of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\xi(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
 & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma r)||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma r}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma s}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
\end{aligned}
\label{eq:main}
\end{equation}

where in transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and 
\begin{equation}
\Phi_{b,\tau}(\boldsymbol{r},\Gamma r')=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
\end{equation}
\end{widetext}
The second to third line of Eq.(\ref{eq:main}) follows from assumption of Gaussian random
sky, and the third to fourth line follows from the assumption that
the 3D power spectrum varies negligibly over the region of interest
so that $\hat{P_{21}}(k+k_{2})\approx\hat{P_{21}}(k)$. Since $\Gamma$
is a sky rotation, it doesn't affect $\nu$, hence we have taken $\nu_{r}$
outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end. This means that correlation is peaked at the same time-offset for all delay channels.

Finally, since the beam pattern and bandwidth are given in $\hat{\boldsymbol{s}}$
and $\nu$, we convert the integral back to these coordinates to get
the general relation between the delay-transformed visibilities and
the power spectrum:

\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\\
 & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma s}\cdot\frac{\boldsymbol{b'}}{c}\right)\right].\end{aligned}
\label{eq:final}
\end{equation}
